On the presentation of orthogonal groups over polynomial rings

### 2019-07-24 dl-paper.tex
Minor fixes

For a triple $(S, P, G)$ we could define the unstable Quillen K-groups functors as $\K_{i}^{G, P}(R) = \pi_i(BG(R)^+_{P(R)})$. The functors with lower numbers admit the following description.
\begin{lemma}\label{lem:lowerKgroups} There are natural isomorphisms \begin{enumerate}
 \item $\K_1^{G,P}(R) \cong G(R) / P(R)$;
 \item $\K_2^{G,P}(R) \cong \Ker(S(R) \to G(R))$;
 \item $\K_3^{G,P}(R) \cong \HH_3(S(R), \ZZ).$ \end{enumerate} \end{lemma}
\begin{proof} The first claim is obvious, the second and the third claim follow from~\ref{req:subfunc} and~\ref{req:uce} using the standard properties of the plus-construction, see \cite[\S~IV.1]{Kbook}   (cf. Exercises~1.8--1.9 ibid.) \end{proof}

\subsection{Some elementary lemmas}
\begin{comment}
Hall-Witt identity
\[ [[ y^{-1}, x], z] ^ {y^{-1}}  [[ z^{-1}, y], x] ^ {z^{-1}}  [[ x^{-1}, z], y] ^ {x^{-1}} = 1 \]
\[ [[ y^{-1}, x], z]  \cdot [[ z^{-1}, y], x] ^ {z^{-1}y} \cdot  [[ x^{-1}, z], y] ^ {x^{-1}y} = 1. \]
\[ [[ z^{-1}, y^{-1}], x] ^ {z^{-1}y^{-1}} \cdot  [[ x^{-1}, z], y^{-1}] ^ {x^{-1}y^{-1}} = [z, [ y, x]] . \]
\[  . \]
\end{comment}
Throughtout this section $\Phi$ is a simply laced root system, $A$ denotes a commutative ring, and $I, J$ are two arbitrary ideals of $A$.

We define the following two families of elements of $\St(\Phi, A)$:
\begin{itemize}
 \item $z_\alpha(s, \xi) := x_\alpha(s)^{x_{-\alpha}(\xi)}$ defined for $\xi \in A$, $s \in I$;
 \item $c_\alpha(s, t) = [x_\alpha(s), x_{-\alpha}(t)]$ defined for $s \in I$, $t \in J$.
\end{itemize}

@@

\begin{rem} For a simply-laced $\Phi$ the second assertion of~\cref{thm:Stepanov} can be deduced from the first one by means of~\cref{Zrels}. Indeed, for a root subset $U\subseteq \Phi$ consider the following operation: \[dU := U \cup (U - U)\cap \Phi.\] In other words, $d$ adjoins to $U$ all differences of roots from $U$ which are themselves roots. It is not hard to show that for any parabolic subset $S \subseteq \Phi$ the subset $\Sigma_S$ has the property that $d^n(\Sigma_S) = \Phi$ for some $n$ (in fact, for $n=2$). It remains to see that relation~\eqref{Z5} immediately implies the following assertion: every group $G$ containing $\mathcal{Z}(U, R, I)$ also contains $\mathcal{Z}(dU, R, I)$. \end{rem}

>>>>

Define the unstable Quillen K-groups functors as $\K_{i}^{G, P}(R) = \pi_i(BG(R)^+_{P(R)})$, $i\geq 1$. The functors with lower numbers admit the following description.
\begin{lemma}\label{lem:lowerKgroups} There are natural isomorphisms \begin{enumerate}
 \item $\K_1^{G,P}(R) \cong G(R) / P(R)$;
 \item $\K_2^{G,P}(R) \cong \Ker(S(R) \to G(R))$;
 \item $\K_3^{G,P}(R) \cong \HH_3(S(R), \ZZ).$ \end{enumerate} \end{lemma}
\begin{proof} The first claim is obvious, the second and the third claim follow from~\ref{req:subfunc} and~\ref{req:uce} using the standard properties of the plus-construction, see \cite[\S~IV.1]{Kbook}   (cf. Exercises~1.8--1.9 ibid.) \end{proof}

\subsection{Elementary calculations in relative Steinberg groups}
\begin{comment}
Hall-Witt identity
\[ [[ y^{-1}, x], z] ^ {y^{-1}}  [[ z^{-1}, y], x] ^ {z^{-1}}  [[ x^{-1}, z], y] ^ {x^{-1}} = 1 \]
\[ [[ y^{-1}, x], z]  \cdot [[ z^{-1}, y], x] ^ {z^{-1}y} \cdot  [[ x^{-1}, z], y] ^ {x^{-1}y} = 1. \]
\[ [[ z^{-1}, y^{-1}], x] ^ {z^{-1}y^{-1}} \cdot  [[ x^{-1}, z], y^{-1}] ^ {x^{-1}y^{-1}} = [z, [ y, x]] . \]
\[  . \]
\end{comment}
Throughtout this section $\Phi$ is a simply laced root system, $A$ is a commutative ring, and $I, J$ are two its ideals.

We denote by $\overline{\St}(\Phi, A, I)$ the kernel of the map $\St(\Phi, A) \to \St(\Phi, A/I)$.
This group coincides with the image in $\St(\Phi, A)$ of the relative group $\St(\Phi, A, I)$ defined in the previous subsection.

We define the following two families of elements of $\overline{\St}(\Phi, A, I)$:
\begin{itemize}
 \item $z_\alpha(s, \xi) := x_\alpha(s)^{x_{-\alpha}(\xi)}$ defined for $\xi \in A$, $s \in I$;
 \item $c_\alpha(s, t) = [x_\alpha(s), x_{-\alpha}(t)]$ defined for $s \in I$, $t \in J$.
\end{itemize}

@@

\begin{rem} For a simply-laced $\Phi$ the second assertion of~\cref{thm:Stepanov} can be deduced from the first one by means of~\cref{Zrels}. Indeed, consider the operator $d \colon 2^\Phi \to 2^\Phi$ of root subsets given by $d(U) = U \cup (U - U)\cap \Phi$, $U \subseteq \Phi$. In other words, $d$ adjoins to $U$ all differences of roots from $U$ which are themselves roots. It is not hard to show that for any parabolic subset $S \subseteq \Phi$ the subset $\Sigma_S$ has the property that $d^2(\Sigma_S) = \Phi$. It remains to see that relation~\eqref{Z5} immediately implies that every group $G$ containing $\mathcal{Z}(U, R, I)$ also contains $\mathcal{Z}(dU, R, I)$. \end{rem}
